\section{Conclusion}

The Personal Financial AI Agent demonstrates that artificial intelligence can effectively augment human financial decision-making. By combining conversational AI, retrieval-augmented generation, and modern web technologies, the system creates an offering that is simultaneously intelligent, accessible, and trustworthy. The modular architecture and emphasis on provider flexibility ensure longevity as the AI landscape continues to evolve. The comprehensive testing and documentation enable both researchers and practitioners to build upon this foundation.

As AI increasingly shapes financial services, maintaining focus on user trust, transparency, and privacy will be crucial. This project provides a template for building AI systems in regulated domains where accuracy and trustworthiness are paramount. Financial advisory is a domain where AI's promise to democratize expertise must be balanced with responsibility to protect users from misinformation and poor recommendations. By emphasizing grounding in real data through RAG, supporting multiple inference options for privacy, and maintaining transparent about limitations, this system demonstrates an approach to trustworthy AI in high-stakes domains.

Despite its strengths, the system has several inherent limitations that should be acknowledged. Model dependency means system quality is fundamentally bounded by underlying LLM capabilities, with smaller models showing limitations in financial reasoning compared to larger models. Data currency presents a challenge since historical financial data has a fixed lookback window, and real-time market data integration would improve recommendations. Limited risk modeling is employed, with simplified volatility metrics; advanced portfolio construction techniques such as mean-variance optimization and factor models could enhance recommendations. The system provides recommendations only and cannot execute trades, so integration with trading platforms would require additional compliance and security measures. Scalability is limited since the current architecture uses in-memory processing, and distributed systems would be needed for enterprise-scale deployment.

The complete source code for the Personal Financial AI Agent is available at \url{https://github.com/merendamattia/personal-financial-ai-agent}.
